---
title: "Data"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(dplyr)
load("USA_county_data.RData")
countydata <- USA_county_data %>% select(fips,name_16,rep16_frac,dem16_frac,rep16_frac2,dem16_frac2,statecode_prev,rep12_frac,dem12_frac,rep12_frac2,dem12_frac2,rep08_frac,dem08_frac,rep08_frac2,dem08_frac2,Less.Than.High.School,At.Least.High.School.Diploma,At.Least.Bachelor.s.Degree,Graduate.Degree,School.Enrollment,Median.Earnings.2010.dollars,White.not.Latino.Population,African.American.Population,Native.American.Population,Asian.American.Population,Latino.Population,Children.Under.6.Living.in.Poverty,Adults.65.and.Older.Living.in.Poverty,Total.Population,Preschool.Enrollment.Ratio.enrolled.ages.3.and.4,Poverty.Rate.below.federal.poverty.threshold,Child.Poverty.living.in.families.below.the.poverty.line,Management.professional.and.related.occupations,Service.occupations,Sales.and.office.occupations,Farming.fishing.and.forestry.occupations,Construction.extraction.maintenance.and.repair.occupations,Production.transportation.and.material.moving.occupations,State,median_age,Poor.physical.health.days,Poor.mental.health.days,Low.birthweight,Teen.births,Children.in.single.parent.households,Adult.smoking,Adult.obesity,Diabetes,Sexually.transmitted.infections,HIV.prevalence.rate,Uninsured,Unemployment)
countydata <- countydata %>% filter(!fips %in% seq(0,100000, by=1000)) %>% arrange(fips)
```


library(readxl)
load("ruralurbancodes2013.xls")
ruralurbancodes2013 <- ruralurbancodes2013 %>% mutate(fips=FIPS) %>% mutate(fips=as.numeric(FIPS))
newcountydata <- full_join(ruralurbancodes2013,countydata,by=c("fips" = "fips"))


```{r}
library(readxl)
read_excel_allsheets <- function(filename) {
    sheets <- readxl::excel_sheets(filename)
    x <-    lapply(sheets, function(X) readxl::read_excel(filename, sheet = X))
    names(x) <- sheets
    x
}
mortalityrates <- read_excel_allsheets("MortalityRatesClean.xlsx")
mortalityrates <- as.data.frame(mortalityrates)
```

```{r}
require(dplyr)
mortalityrates <- rename(mortalityrates, fips = HIV.FIPS)
mortalityrates <- rename(mortalityrates, location = HIV.Location)
mortalityrates <- mortalityrates %>% select(location,fips,HIV.HIV80,HIV.HIV85,HIV.HIV90,HIV.HIV95,HIV.HIV00,HIV.HIV05,HIV.HIV10,HIV.HIV14,INF.INF80,INF.INF85,INF.INF90,INF.INF95,INF.INF00,INF.INF05,INF.INF10,INF.INF14,TROP.TROP80,TROP.TROP85,TROP.TROP90,TROP.TROP95,TROP.TROP00,TROP.TROP05,TROP.TROP10,TROP.TROP14,MAT.MAT80,MAT.MAT85,MAT.MAT90,MAT.MAT95,MAT.MAT00,MAT.MAT05,MAT.MAT10,MAT.MAT14,NEON.NEON80,NEON.NEON85,NEON.NEON90,NEON.NEON95,NEON.NEON00,NEON.NEON05,NEON.NEON10,NEON.NEON14,NUT.NUT80,NUT.NUT85,NUT.NUT90,NUT.NUT95,NUT.NUT00,NUT.NUT05,NUT.NUT10,NUT.NUT14,OTHC.OTHC80,OTHC.OTHC85,OTHC.OTHC90,OTHC.OTHC95,OTHC.OTHC00,OTHC.OTHC05,OTHC.OTHC10,OTHC.OTHC14,NEOP.NEOP80,NEOP.NEOP85,NEOP.NEOP90,NEOP.NEOP95,NEOP.NEOP00,NEOP.NEOP05,NEOP.NEOP10,NEOP.NEOP14,CHRON.CHRON80,CHRON.CHRON85,CHRON.CHRON90,CHRON.CHRON95,CHRON.CHRON00,CHRON.CHRON05,CHRON.CHRON10,CHRON.CHRON14,CIRR.CIRR80,CIRR.CIRR85,CIRR.CIRR90,CIRR.CIRR95,CIRR.CIRR00,CIRR.CIRR05,CIRR.CIRR10,CIRR.CIRR14,DIG.DIG80,DIG.DIG85,DIG.DIG90,DIG.DIG95,DIG.DIG00,DIG.DIG05,DIG.DIG10,DIG.DIG14,NEUR.NEUR80,NEUR.NEUR85,NEUR.NEUR90,NEUR.NEUR95,NEUR.NEUR00,NEUR.NEUR05,NEUR.NEUR10,NEUR.NEUR14,MENSUB.MENSUB80,MENSUB.MENSUB85,MENSUB.MENSUB90,MENSUB.MENSUB95,MENSUB.MENSUB00,MENSUB.MENSUB05,MENSUB.MENSUB10,MENSUB.MENSUB14,DIAB.DIAB80,DIAB.DIAB85,DIAB.DIAB90,DIAB.DIAB95,DIAB.DIAB00,DIAB.DIAB05,DIAB.DIAB10,DIAB.DIAB14,MUSC.MUSC80,MUSC.MUSC85,MUSC.MUSC90,MUSC.MUSC95,MUSC.MUSC00,MUSC.MUSC05,MUSC.MUSC10,MUSC.MUSC14,OTHN.OTHN80,OTHN.OTHN85,OTHN.OTHN90,OTHN.OTHN95,OTHN.OTHN00,OTHN.OTHN05,OTHN.OTHN10,OTHN.OTHN14,TRAN.TRAN80,TRAN.TRAN85,TRAN.TRAN90,TRAN.TRAN95,TRAN.TRAN00,TRAN.TRAN05,TRAN.TRAN10,TRAN.TRAN14,UNIN.UNIN80,UNIN.UNIN85,UNIN.UNIN90,UNIN.UNIN95,UNIN.UNIN00,UNIN.UNIN05,UNIN.UNIN10,UNIN.UNIN14,SELF.SELF80,SELF.SELF85,SELF.SELF90,SELF.SELF95,SELF.SELF00,SELF.SELF05,SELF.SELF10,SELF.SELF14,WAR.WAR80,WAR.WAR85,WAR.WAR90,WAR.WAR95,WAR.WAR00,WAR.WAR05,WAR.WAR10,WAR.WAR14)
```

HIV: HIV-AIDS and Tuberculosis
INF: Diarrhea, lower respiratory, and other common infectious diseases
TROP: Neglected tropical diseases and malaria
MAT: Maternal disorders
NEON: Neonatal disorders
NUT: Nutritional deficiencies
OTHC: Other communicable, maternal, neonatal, and nutritional diseases
NEOP: Neoplasms
CHRON: Chronic respiratory diseases
CIRR: Cirrhosis and other chronic liver diseases
DIG: Digestive diseases
NEUR: Neurological disorders
MENSUB: Mental and substance use disorders
DIAB: Diabetes, urogenital, blood, and endocrine diseases
MUSC: Musculoskeletal disorders
OTHN: Other non-communicable diseases
TRAN: Transport injuries
UNIN: Unintentional injuries
SELF: Self-harm and interpersonal violence
WAR: Forces of nature, war, and legal intervention

```{r}
mortalityrates <- mortalityrates %>% filter(fips>100) %>% filter(fips<70000) %>% arrange(fips)

require(dplyr)
countyrates <- inner_join(newcountydata,mortalityrates,by=c("fips"))
```

```{r}
countyratesnogaps <- countyrates %>% filter(fips>100) %>% filter(fips<70000) %>% filter(!is.na(rep16_frac)) %>% filter(!is.na(At.Least.High.School.Diploma))  %>% arrange(fips)
```

```{r}
countyratesnogaps <- countyratesnogaps %>% select(location,fips,HIV.HIV80,HIV.HIV85,HIV.HIV90,HIV.HIV95,HIV.HIV00,HIV.HIV05,HIV.HIV10,HIV.HIV14,INF.INF80,INF.INF85,INF.INF90,INF.INF95,INF.INF00,INF.INF05,INF.INF10,INF.INF14,TROP.TROP80,TROP.TROP85,TROP.TROP90,TROP.TROP95,TROP.TROP00,TROP.TROP05,TROP.TROP10,TROP.TROP14,MAT.MAT80,MAT.MAT85,MAT.MAT90,MAT.MAT95,MAT.MAT00,MAT.MAT05,MAT.MAT10,MAT.MAT14,NEON.NEON80,NEON.NEON85,NEON.NEON90,NEON.NEON95,NEON.NEON00,NEON.NEON05,NEON.NEON10,NEON.NEON14,NUT.NUT80,NUT.NUT85,NUT.NUT90,NUT.NUT95,NUT.NUT00,NUT.NUT05,NUT.NUT10,NUT.NUT14,OTHC.OTHC80,OTHC.OTHC85,OTHC.OTHC90,OTHC.OTHC95,OTHC.OTHC00,OTHC.OTHC05,OTHC.OTHC10,OTHC.OTHC14,NEOP.NEOP80,NEOP.NEOP85,NEOP.NEOP90,NEOP.NEOP95,NEOP.NEOP00,NEOP.NEOP05,NEOP.NEOP10,NEOP.NEOP14,CHRON.CHRON80,CHRON.CHRON85,CHRON.CHRON90,CHRON.CHRON95,CHRON.CHRON00,CHRON.CHRON05,CHRON.CHRON10,CHRON.CHRON14,CIRR.CIRR80,CIRR.CIRR85,CIRR.CIRR90,CIRR.CIRR95,CIRR.CIRR00,CIRR.CIRR05,CIRR.CIRR10,CIRR.CIRR14,DIG.DIG80,DIG.DIG85,DIG.DIG90,DIG.DIG95,DIG.DIG00,DIG.DIG05,DIG.DIG10,DIG.DIG14,NEUR.NEUR80,NEUR.NEUR85,NEUR.NEUR90,NEUR.NEUR95,NEUR.NEUR00,NEUR.NEUR05,NEUR.NEUR10,NEUR.NEUR14,MENSUB.MENSUB80,MENSUB.MENSUB85,MENSUB.MENSUB90,MENSUB.MENSUB95,MENSUB.MENSUB00,MENSUB.MENSUB05,MENSUB.MENSUB10,MENSUB.MENSUB14,DIAB.DIAB80,DIAB.DIAB85,DIAB.DIAB90,DIAB.DIAB95,DIAB.DIAB00,DIAB.DIAB05,DIAB.DIAB10,DIAB.DIAB14,MUSC.MUSC80,MUSC.MUSC85,MUSC.MUSC90,MUSC.MUSC95,MUSC.MUSC00,MUSC.MUSC05,MUSC.MUSC10,MUSC.MUSC14,OTHN.OTHN80,OTHN.OTHN85,OTHN.OTHN90,OTHN.OTHN95,OTHN.OTHN00,OTHN.OTHN05,OTHN.OTHN10,OTHN.OTHN14,TRAN.TRAN80,TRAN.TRAN85,TRAN.TRAN90,TRAN.TRAN95,TRAN.TRAN00,TRAN.TRAN05,TRAN.TRAN10,TRAN.TRAN14,UNIN.UNIN80,UNIN.UNIN85,UNIN.UNIN90,UNIN.UNIN95,UNIN.UNIN00,UNIN.UNIN05,UNIN.UNIN10,UNIN.UNIN14,SELF.SELF80,SELF.SELF85,SELF.SELF90,SELF.SELF95,SELF.SELF00,SELF.SELF05,SELF.SELF10,SELF.SELF14,WAR.WAR80,WAR.WAR85,WAR.WAR90,WAR.WAR95,WAR.WAR00,WAR.WAR05,WAR.WAR10,WAR.WAR14,rep16_frac,dem16_frac,Less.Than.High.School,At.Least.High.School.Diploma,At.Least.Bachelor.s.Degree,Graduate.Degree,School.Enrollment,Median.Earnings.2010.dollars,White.not.Latino.Population,African.American.Population,Native.American.Population,Asian.American.Population,Latino.Population,Adults.65.and.Older.Living.in.Poverty,Total.Population,Poverty.Rate.below.federal.poverty.threshold,Child.Poverty.living.in.families.below.the.poverty.line,Management.professional.and.related.occupations,Service.occupations,Sales.and.office.occupations,Farming.fishing.and.forestry.occupations,Construction.extraction.maintenance.and.repair.occupations,Production.transportation.and.material.moving.occupations,median_age,Children.in.single.parent.households,Adult.obesity,Diabetes,Uninsured,Unemployment)
countyratesfullcases <- countyratesnogaps[complete.cases(countyratesnogaps),]
```

```{r}
require(caret)
require(rpart)
require(rpart.plot)
library(tree)
```

HIV: HIV-AIDS and Tuberculosis

```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.HIV14$finalModel
rpart.plot(tr.HIV14$finalModel)
preds.tr.HIV14 <- predict(tr.HIV14$finalModel, newdata = countyratesfullcases)
realHIV14 <- countyratesfullcases$HIV.HIV14
HIV14 <- cbind.data.frame(preds.tr.HIV14,realHIV14)
HIV14 <- HIV14 %>% mutate(HIVtreediff = ((preds.tr.HIV14 - realHIV14)/realHIV14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.HIV14 <- predict(knn.tr.HIV14, newdata = countyratesfullcases)
realknnHIV14 <- countyratesfullcases$HIV.HIV14
knnHIV14 <- cbind.data.frame(preds.knn.tr.HIV14,realknnHIV14)
knnHIV14 <- knnHIV14 %>% mutate(HIVknndiff = ((preds.knn.tr.HIV14 - realknnHIV14)/realknnHIV14))
```


INF: Diarrhea, lower respiratory, and other common infectious diseases

```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.INF14 <- train(INF.INF14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.INF14$finalModel
rpart.plot(tr.INF14$finalModel)
preds.tr.INF14 <- predict(tr.INF14$finalModel, newdata = countyratesfullcases)
realINF14 <- countyratesfullcases$INF.INF14
INF14 <- cbind.data.frame(preds.tr.INF14,realINF14)
INF14 <- INF14 %>% mutate(INFtreediff = ((preds.tr.INF14 - realINF14)/realINF14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.INF14 <- train(INF.INF14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.INF14 <- predict(knn.tr.INF14, newdata = countyratesfullcases)
realknnINF14 <- countyratesfullcases$INF.INF14
knnINF14 <- cbind.data.frame(preds.knn.tr.INF14,realknnINF14)
knnINF14 <- knnINF14 %>% mutate(INFknndiff = ((preds.knn.tr.INF14 - realknnINF14)/realknnINF14))
```


TROP: Neglected tropical diseases and malaria


```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.TROP14 <- train(TROP.TROP14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.TROP14$finalModel
rpart.plot(tr.TROP14$finalModel)
preds.tr.TROP14 <- predict(tr.TROP14$finalModel, newdata = countyratesfullcases)
realTROP14 <- countyratesfullcases$TROP.TROP14
TROP14 <- cbind.data.frame(preds.tr.TROP14,realTROP14)
TROP14 <- TROP14 %>% mutate(TROPtreediff = ((preds.tr.TROP14 - realTROP14)/realTROP14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.TROP14 <- train(TROP.TROP14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.TROP14 <- predict(knn.tr.TROP14, newdata = countyratesfullcases)
realknnTROP14 <- countyratesfullcases$TROP.TROP14
knnTROP14 <- cbind.data.frame(preds.knn.tr.TROP14,realknnTROP14)
knnTROP14 <- knnTROP14 %>% mutate(TROPknndiff = ((preds.knn.tr.TROP14 - realknnTROP14)/realknnTROP14))
```


MAT: Maternal disorders

```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.MAT14 <- train(MAT.MAT14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.MAT14$finalModel
rpart.plot(tr.MAT14$finalModel)
preds.tr.MAT14 <- predict(tr.MAT14$finalModel, newdata = countyratesfullcases)
realMAT14 <- countyratesfullcases$MAT.MAT14
MAT14 <- cbind.data.frame(preds.tr.MAT14,realMAT14)
MAT14 <- MAT14 %>% mutate(MATtreediff = ((preds.tr.MAT14 - realMAT14)/realMAT14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.MAT14 <- train(MAT.MAT14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.MAT14 <- predict(knn.tr.MAT14, newdata = countyratesfullcases)
realknnMAT14 <- countyratesfullcases$MAT.MAT14
knnMAT14 <- cbind.data.frame(preds.knn.tr.MAT14,realknnMAT14)
knnMAT14 <- knnMAT14 %>% mutate(MATknndiff = ((preds.knn.tr.MAT14 - realknnMAT14)/realknnMAT14))
```


NEON: Neonatal disorders


```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.NEON14 <- train(NEON.NEON14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.NEON14$finalModel
rpart.plot(tr.NEON14$finalModel)
preds.tr.NEON14 <- predict(tr.NEON14$finalModel, newdata = countyratesfullcases)
realNEON14 <- countyratesfullcases$NEON.NEON14
NEON14 <- cbind.data.frame(preds.tr.NEON14,realNEON14)
NEON14 <- NEON14 %>% mutate(NEONtreediff = ((preds.tr.NEON14 - realNEON14)/realNEON14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.NEON14 <- train(NEON.NEON14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.NEON14 <- predict(knn.tr.NEON14, newdata = countyratesfullcases)
realknnNEON14 <- countyratesfullcases$NEON.NEON14
knnNEON14 <- cbind.data.frame(preds.knn.tr.NEON14,realknnNEON14)
knnNEON14 <- knnNEON14 %>% mutate(NEONknndiff = ((preds.knn.tr.NEON14 - realknnNEON14)/realknnNEON14))
```


NUT: Nutritional deficiencies


```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.NUT14 <- train(NUT.NUT14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.NUT14$finalModel
rpart.plot(tr.NUT14$finalModel)
preds.tr.NUT14 <- predict(tr.NUT14$finalModel, newdata = countyratesfullcases)
realNUT14 <- countyratesfullcases$NUT.NUT14
NUT14 <- cbind.data.frame(preds.tr.NUT14,realNUT14)
NUT14 <- NUT14 %>% mutate(NUTtreediff = ((preds.tr.NUT14 - realNUT14)/realNUT14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.NUT14 <- train(NUT.NUT14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.NUT14 <- predict(knn.tr.NUT14, newdata = countyratesfullcases)
realknnNUT14 <- countyratesfullcases$NUT.NUT14
knnNUT14 <- cbind.data.frame(preds.knn.tr.NUT14,realknnNUT14)
knnNUT14 <- knnHIV14 %>% mutate(NUTknndiff = ((preds.knn.tr.NUT14 - realknnNUT14)/realknnNUT14))
```


OTHC: Other communicable, maternal, neonatal, and nutritional diseases


```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.OTHC14 <- train(OTHC.OTHC14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.OTHC14$finalModel
rpart.plot(tr.OTHC14$finalModel)
preds.tr.OTHC14 <- predict(tr.OTHC14$finalModel, newdata = countyratesfullcases)
realOTHC14 <- countyratesfullcases$OTHC.OTHC14
OTHC14 <- cbind.data.frame(preds.tr.OTHC14,realOTHC14)
OTHC14 <- OTHC14 %>% mutate(OTHCtreediff = ((preds.tr.OTHC14 - realOTHC14)/realOTHC14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.OTHC14 <- train(OTHC.OTHC14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.OTHC14 <- predict(knn.tr.OTHC14, newdata = countyratesfullcases)
realknnOTHC14 <- countyratesfullcases$OTHC.OTHC14
knnOTHC14 <- cbind.data.frame(preds.knn.tr.OTHC14,realknnOTHC14)
knnOTHC14 <- knnOTHC14 %>% mutate(OTHCknndiff = ((preds.knn.tr.OTHC14 - realknnOTHC14)/realknnOTHC14))
```


NEOP: Neoplasms


```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.NEOP14 <- train(NEOP.NEOP14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.NEOP14$finalModel
rpart.plot(tr.NEOP14$finalModel)
preds.tr.NEOP14 <- predict(tr.NEOP14$finalModel, newdata = countyratesfullcases)
realNEOP14 <- countyratesfullcases$NEOP.NEOP14
NEOP14 <- cbind.data.frame(preds.tr.NEOP14,realNEOP14)
NEOP14 <- NEOP14 %>% mutate(NEOPtreediff = ((preds.tr.NEOP14 - realNEOP14)/realNEOP14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.NEOP14 <- train(NEOP.NEOP14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.NEOP14 <- predict(knn.tr.NEOP14, newdata = countyratesfullcases)
realknnNEOP14 <- countyratesfullcases$NEOP.NEOP14
knnNEOP14 <- cbind.data.frame(preds.knn.tr.NEOP14,realknnNEOP14)
knnNEOP14 <- knnNEOP14 %>% mutate(NEOPknndiff = ((preds.knn.tr.NEOP14 - realknnNEOP14)/realknnNEOP14))
```


```{r}
fips <- countyratesfullcases$fips
results  <- cbind.data.frame(fips,HIV14$realHIV14,HIV14$HIVtreediff,knnHIV14$HIVknndiff,INF14$realINF14,INF14$INFtreediff,knnINF14$INFknndiff,TROP14$realTROP14,TROP14$TROPtreediff,knnTROP14$TROPknndiff,MAT14$realMAT14,MAT14$MATtreediff,knnMAT14$MATknndiff,NEON14$realNEON14,NEON14$NEONtreediff,knnNEON14$NEONknndiff,NUT14$realNUT14,NUT14$NUTtreediff,knnNUT14$NUTknndiff,OTHC14$realOTHC14,OTHC14$OTHCtreediff,knnOTHC14$OTHCknndiff,NEOP14$realNEOP14,NEOP14$NEOPtreediff,knnNEOP14$NEOPknndiff)
View(results)

```


```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.HIV14$finalModel
rpart.plot(tr.HIV14$finalModel)
preds.tr.HIV14 <- predict(tr.HIV14$finalModel, newdata = countyratesfullcases)
realHIV14 <- countyratesfullcases$HIV.HIV14
HIV14 <- cbind.data.frame(preds.tr.HIV14,realHIV14)
HIV14 <- HIV14 %>% mutate(HIVtreediff = ((preds.tr.HIV14 - realHIV14)/realHIV14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.HIV14 <- predict(knn.tr.HIV14, newdata = countyratesfullcases)
realknnHIV14 <- countyratesfullcases$HIV.HIV14
knnHIV14 <- cbind.data.frame(preds.knn.tr.HIV14,realknnHIV14)
knnHIV14 <- knnHIV14 %>% mutate(HIVknndiff = ((preds.knn.tr.HIV14 - realknnHIV14)/realknnHIV14))
```


```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.HIV14$finalModel
rpart.plot(tr.HIV14$finalModel)
preds.tr.HIV14 <- predict(tr.HIV14$finalModel, newdata = countyratesfullcases)
realHIV14 <- countyratesfullcases$HIV.HIV14
HIV14 <- cbind.data.frame(preds.tr.HIV14,realHIV14)
HIV14 <- HIV14 %>% mutate(HIVtreediff = ((preds.tr.HIV14 - realHIV14)/realHIV14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.HIV14 <- predict(knn.tr.HIV14, newdata = countyratesfullcases)
realknnHIV14 <- countyratesfullcases$HIV.HIV14
knnHIV14 <- cbind.data.frame(preds.knn.tr.HIV14,realknnHIV14)
knnHIV14 <- knnHIV14 %>% mutate(HIVknndiff = ((preds.knn.tr.HIV14 - realknnHIV14)/realknnHIV14))
```


```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.HIV14$finalModel
rpart.plot(tr.HIV14$finalModel)
preds.tr.HIV14 <- predict(tr.HIV14$finalModel, newdata = countyratesfullcases)
realHIV14 <- countyratesfullcases$HIV.HIV14
HIV14 <- cbind.data.frame(preds.tr.HIV14,realHIV14)
HIV14 <- HIV14 %>% mutate(HIVtreediff = ((preds.tr.HIV14 - realHIV14)/realHIV14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.HIV14 <- predict(knn.tr.HIV14, newdata = countyratesfullcases)
realknnHIV14 <- countyratesfullcases$HIV.HIV14
knnHIV14 <- cbind.data.frame(preds.knn.tr.HIV14,realknnHIV14)
knnHIV14 <- knnHIV14 %>% mutate(HIVknndiff = ((preds.knn.tr.HIV14 - realknnHIV14)/realknnHIV14))
```


```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.HIV14$finalModel
rpart.plot(tr.HIV14$finalModel)
preds.tr.HIV14 <- predict(tr.HIV14$finalModel, newdata = countyratesfullcases)
realHIV14 <- countyratesfullcases$HIV.HIV14
HIV14 <- cbind.data.frame(preds.tr.HIV14,realHIV14)
HIV14 <- HIV14 %>% mutate(HIVtreediff = ((preds.tr.HIV14 - realHIV14)/realHIV14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.HIV14 <- predict(knn.tr.HIV14, newdata = countyratesfullcases)
realknnHIV14 <- countyratesfullcases$HIV.HIV14
knnHIV14 <- cbind.data.frame(preds.knn.tr.HIV14,realknnHIV14)
knnHIV14 <- knnHIV14 %>% mutate(HIVknndiff = ((preds.knn.tr.HIV14 - realknnHIV14)/realknnHIV14))
```


```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.HIV14$finalModel
rpart.plot(tr.HIV14$finalModel)
preds.tr.HIV14 <- predict(tr.HIV14$finalModel, newdata = countyratesfullcases)
realHIV14 <- countyratesfullcases$HIV.HIV14
HIV14 <- cbind.data.frame(preds.tr.HIV14,realHIV14)
HIV14 <- HIV14 %>% mutate(HIVtreediff = ((preds.tr.HIV14 - realHIV14)/realHIV14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.HIV14 <- predict(knn.tr.HIV14, newdata = countyratesfullcases)
realknnHIV14 <- countyratesfullcases$HIV.HIV14
knnHIV14 <- cbind.data.frame(preds.knn.tr.HIV14,realknnHIV14)
knnHIV14 <- knnHIV14 %>% mutate(HIVknndiff = ((preds.knn.tr.HIV14 - realknnHIV14)/realknnHIV14))
```


```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.HIV14$finalModel
rpart.plot(tr.HIV14$finalModel)
preds.tr.HIV14 <- predict(tr.HIV14$finalModel, newdata = countyratesfullcases)
realHIV14 <- countyratesfullcases$HIV.HIV14
HIV14 <- cbind.data.frame(preds.tr.HIV14,realHIV14)
HIV14 <- HIV14 %>% mutate(HIVtreediff = ((preds.tr.HIV14 - realHIV14)/realHIV14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.HIV14 <- predict(knn.tr.HIV14, newdata = countyratesfullcases)
realknnHIV14 <- countyratesfullcases$HIV.HIV14
knnHIV14 <- cbind.data.frame(preds.knn.tr.HIV14,realknnHIV14)
knnHIV14 <- knnHIV14 %>% mutate(HIVknndiff = ((preds.knn.tr.HIV14 - realknnHIV14)/realknnHIV14))
```


```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.HIV14$finalModel
rpart.plot(tr.HIV14$finalModel)
preds.tr.HIV14 <- predict(tr.HIV14$finalModel, newdata = countyratesfullcases)
realHIV14 <- countyratesfullcases$HIV.HIV14
HIV14 <- cbind.data.frame(preds.tr.HIV14,realHIV14)
HIV14 <- HIV14 %>% mutate(HIVtreediff = ((preds.tr.HIV14 - realHIV14)/realHIV14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.HIV14 <- predict(knn.tr.HIV14, newdata = countyratesfullcases)
realknnHIV14 <- countyratesfullcases$HIV.HIV14
knnHIV14 <- cbind.data.frame(preds.knn.tr.HIV14,realknnHIV14)
knnHIV14 <- knnHIV14 %>% mutate(HIVknndiff = ((preds.knn.tr.HIV14 - realknnHIV14)/realknnHIV14))
```


```{r trees}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
tr.HIV14$finalModel
rpart.plot(tr.HIV14$finalModel)
preds.tr.HIV14 <- predict(tr.HIV14$finalModel, newdata = countyratesfullcases)
realHIV14 <- countyratesfullcases$HIV.HIV14
HIV14 <- cbind.data.frame(preds.tr.HIV14,realHIV14)
HIV14 <- HIV14 %>% mutate(HIVtreediff = ((preds.tr.HIV14 - realHIV14)/realHIV14))
```

```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.HIV14 <- predict(knn.tr.HIV14, newdata = countyratesfullcases)
realknnHIV14 <- countyratesfullcases$HIV.HIV14
knnHIV14 <- cbind.data.frame(preds.knn.tr.HIV14,realknnHIV14)
knnHIV14 <- knnHIV14 %>% mutate(HIVknndiff = ((preds.knn.tr.HIV14 - realknnHIV14)/realknnHIV14))
```