---
title: "Death by Chocolate? No, Death by County. Modelling County-Level Mortality Rates"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=60),tidy=TRUE)
```


**Introduction**

Mortality rates from many causes of death are known to vary across counties. We wanted to explore what other factors about counties are correlated to county-level mortality rates from various causes to better understand what some causes of the relative geographic prevalence of different causes of death may be. Our analysis and visualizations display actual variation in causes of death between counties, and show both the power and the shortcomings of using statistical modelling techniques to estimate causes of death based on other factors about counties.



**Motivation/Purpose**

It is understood that mortality rates from different causes vary between regions of the United States and even between counties in the same state, but those differences and potential reasons for them have not been systematically explored.

**Background**

\newpage

**Data Sources**

The first data source we used was [a compilation of county-level data](https://github.com/Deleetdk/USA.county.data/tree/master/data) from a variety of sources compiled by Github user Deleetdk. Most of the data sources included are those from [a paper](https://openpsych.net/files/papers/Kirkegaard_2016e.pdf) published by Emil O. W. Kirkegaard in the journal *Open Quantitative Sociology and Political Science* in 2016, "Inequality across US counties: an S factor analysis". This data source was, very luckily for us, an RData file on GitHub that was very easy to download.

The demographic data (such as racial composition, gender, and age) come from the United States Census Bureau's annual American Community Survey through the [American FactFinder](https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?src=bkmk). This is also the source for some economic indicators (educational attainment, sector composition of employment).
Another data source is the [American Human Development Index of the Measure of America](http://www.measureofamerica.org/project/), which contains health, education, and income indicators.
Many health indicators come from [the Robert Wood Johnson Foundation's County Health Rankings & Roadmaps](http://www.countyhealthrankings.org/explore-health-rankings/rankings-data).
Finally, the county-level electoral data comes from the [failing New York Times](https://www.nytimes.com/elections/results/president).

The first step of data wrangling with this dataset was selecting the variables we wanted from it. We then removed the state-level rows, which all have FIPS codes of X000, so that the data was only counties.

```{r}
require(dplyr)
load("USA_county_data.RData")
countydata <- USA_county_data %>% select(fips,name_16,rep16_frac,dem16_frac,rep16_frac2,dem16_frac2,statecode_prev,rep12_frac,dem12_frac,rep12_frac2,dem12_frac2,rep08_frac,dem08_frac,rep08_frac2,dem08_frac2,Less.Than.High.School,At.Least.High.School.Diploma,At.Least.Bachelor.s.Degree,Graduate.Degree,School.Enrollment,Median.Earnings.2010.dollars,White.not.Latino.Population,African.American.Population,Native.American.Population,Asian.American.Population,Latino.Population,Children.Under.6.Living.in.Poverty,Adults.65.and.Older.Living.in.Poverty,Total.Population,Preschool.Enrollment.Ratio.enrolled.ages.3.and.4,Poverty.Rate.below.federal.poverty.threshold,Child.Poverty.living.in.families.below.the.poverty.line,Management.professional.and.related.occupations,Service.occupations,Sales.and.office.occupations,Farming.fishing.and.forestry.occupations,Construction.extraction.maintenance.and.repair.occupations,Production.transportation.and.material.moving.occupations,State,median_age,Poor.physical.health.days,Poor.mental.health.days,Low.birthweight,Teen.births,Children.in.single.parent.households,Adult.smoking,Adult.obesity,Diabetes,Sexually.transmitted.infections,HIV.prevalence.rate,Uninsured,Unemployment)

countydata <- countydata %>% filter(!fips %in% seq(0,100000, by=1000)) %>% arrange(fips)
```
 

The key dataset for this analysis is a set of estimates of “annual mortality rates by US county for 21 mutually exclusive causes of death from 1980 through 2014” constructed by Laura Dwyer-Lindgren, Amelia Bertozzi-Villa, Rebecca Stubbs, et al. from the National Center for Health Statistics’ National Vital Statistics System and published in the [Journal of the American Medical Association](https://jamanetwork.com/journals/jama/fullarticle/2592499). We downloaded the dataset itself from [Kaggle](https://www.kaggle.com/IHME/us-countylevel-mortality), which it was provided to by the Institute for Health Metrics and Evaluation.

The causes of death (and our short names for them) are:

HIV: HIV-AIDS and Tuberculosis

INF: Diarrhea, lower respiratory, and other common infectious diseases

TROP: Neglected tropical diseases and malaria

MAT: Maternal disorders

NEON: Neonatal disorders

NUT: Nutritional deficiencies

OTHC: Other communicable, maternal, neonatal, and nutritional diseases

NEOP: Neoplasms

CHRON: Chronic respiratory diseases

CIRR: Cirrhosis and other chronic liver diseases

DIG: Digestive diseases

NEUR: Neurological disorders

MENSUB: Mental and substance use disorders

DIAB: Diabetes, urogenital, blood, and endocrine diseases

MUSC: Musculoskeletal disorders

OTHN: Other non-communicable diseases

TRAN: Transport injuries

UNIN: Unintentional injuries

SELF: Self-harm and interpersonal violence

WAR: Forces of nature, war, and legal intervention

This dataset was an Excel file with estimates with estimated mortality rates for each county for each of 20 causes of mortality for each of many years (1980, 1985, 1990, 1995, 2000, 2005, 2010, and 2014). The first thing we had to do was change each estimate from a midpoint and a confidence interval to just the midpoint. We also renamed the sheets and columns to make them easier to work with.
Then, because each cause of death was its own sheet, we had to make a function to read in all of the sheets to one dataframe in RStudio.

```{r}
library(readxl)
read_excel_allsheets <- function(filename) {
    sheets <- readxl::excel_sheets(filename)
    x <-    lapply(sheets, function(X) readxl::read_excel(filename, sheet = X))
    names(x) <- sheets
    x
}
mortalityrates <- read_excel_allsheets("MortalityRatesClean.xlsx")
mortalityrates <- as.data.frame(mortalityrates)
```


Because each cause of death was its own sheet, when we imported the data there were 20 columns of the FIPS codes and 20 columns of location (the county names). As a result, we selected one of the location and FIPS and all of the mortality rate estimates to keep.

```{r}
require(dplyr)
mortalityrates <- rename(mortalityrates, fips = HIV.FIPS)
mortalityrates <- rename(mortalityrates, location = HIV.Location)
mortalityrates <- mortalityrates %>% select(location,fips,HIV.HIV80,HIV.HIV85,HIV.HIV90,HIV.HIV95,HIV.HIV00,HIV.HIV05,HIV.HIV10,HIV.HIV14,INF.INF80,INF.INF85,INF.INF90,INF.INF95,INF.INF00,INF.INF05,INF.INF10,INF.INF14,TROP.TROP80,TROP.TROP85,TROP.TROP90,TROP.TROP95,TROP.TROP00,TROP.TROP05,TROP.TROP10,TROP.TROP14,MAT.MAT80,MAT.MAT85,MAT.MAT90,MAT.MAT95,MAT.MAT00,MAT.MAT05,MAT.MAT10,MAT.MAT14,NEON.NEON80,NEON.NEON85,NEON.NEON90,NEON.NEON95,NEON.NEON00,NEON.NEON05,NEON.NEON10,NEON.NEON14,NUT.NUT80,NUT.NUT85,NUT.NUT90,NUT.NUT95,NUT.NUT00,NUT.NUT05,NUT.NUT10,NUT.NUT14,OTHC.OTHC80,OTHC.OTHC85,OTHC.OTHC90,OTHC.OTHC95,OTHC.OTHC00,OTHC.OTHC05,OTHC.OTHC10,OTHC.OTHC14,NEOP.NEOP80,NEOP.NEOP85,NEOP.NEOP90,NEOP.NEOP95,NEOP.NEOP00,NEOP.NEOP05,NEOP.NEOP10,NEOP.NEOP14,CHRON.CHRON80,CHRON.CHRON85,CHRON.CHRON90,CHRON.CHRON95,CHRON.CHRON00,CHRON.CHRON05,CHRON.CHRON10,CHRON.CHRON14,CIRR.CIRR80,CIRR.CIRR85,CIRR.CIRR90,CIRR.CIRR95,CIRR.CIRR00,CIRR.CIRR05,CIRR.CIRR10,CIRR.CIRR14,DIG.DIG80,DIG.DIG85,DIG.DIG90,DIG.DIG95,DIG.DIG00,DIG.DIG05,DIG.DIG10,DIG.DIG14,NEUR.NEUR80,NEUR.NEUR85,NEUR.NEUR90,NEUR.NEUR95,NEUR.NEUR00,NEUR.NEUR05,NEUR.NEUR10,NEUR.NEUR14,MENSUB.MENSUB80,MENSUB.MENSUB85,MENSUB.MENSUB90,MENSUB.MENSUB95,MENSUB.MENSUB00,MENSUB.MENSUB05,MENSUB.MENSUB10,MENSUB.MENSUB14,DIAB.DIAB80,DIAB.DIAB85,DIAB.DIAB90,DIAB.DIAB95,DIAB.DIAB00,DIAB.DIAB05,DIAB.DIAB10,DIAB.DIAB14,MUSC.MUSC80,MUSC.MUSC85,MUSC.MUSC90,MUSC.MUSC95,MUSC.MUSC00,MUSC.MUSC05,MUSC.MUSC10,MUSC.MUSC14,OTHN.OTHN80,OTHN.OTHN85,OTHN.OTHN90,OTHN.OTHN95,OTHN.OTHN00,OTHN.OTHN05,OTHN.OTHN10,OTHN.OTHN14,TRAN.TRAN80,TRAN.TRAN85,TRAN.TRAN90,TRAN.TRAN95,TRAN.TRAN00,TRAN.TRAN05,TRAN.TRAN10,TRAN.TRAN14,UNIN.UNIN80,UNIN.UNIN85,UNIN.UNIN90,UNIN.UNIN95,UNIN.UNIN00,UNIN.UNIN05,UNIN.UNIN10,UNIN.UNIN14,SELF.SELF80,SELF.SELF85,SELF.SELF90,SELF.SELF95,SELF.SELF00,SELF.SELF05,SELF.SELF10,SELF.SELF14,WAR.WAR80,WAR.WAR85,WAR.WAR90,WAR.WAR95,WAR.WAR00,WAR.WAR05,WAR.WAR10,WAR.WAR14)
```

Similar to with the other dataset, there were state rows as well as county rows. However, in this dataset the states had low FIPS codes (below 100) instead of FIPS codes of multiples of 1000. Additionally, the mortality rates dataset had counties in Puerto Rico (FIPS codes in the 72000s), but we removed these because they did not have data in the Deleetdk data.

```{r}
mortalityrates <- mortalityrates %>% filter(fips>100) %>% filter(fips<70000) %>% arrange(fips)
```


At this point, we combined the two datasets.

```{r}
require(dplyr)
countyrates <- inner_join(countydata,mortalityrates,by=c("fips"))
```

When we thought we needed the latitude and longitude of each county, we downloaded a file that contained that information and renamed and selected the variables for them.

```{r}
require(dplyr)
`2015_Gaz_counties_national` <- read.delim("2015_Gaz_counties_national.txt")
`2015_Gaz_counties_national` <- rename(`2015_Gaz_counties_national`, fips = GEOID)
`2015_Gaz_counties_national` <- rename(`2015_Gaz_counties_national`, latitude = INTPTLAT)
`2015_Gaz_counties_national` <- rename(`2015_Gaz_counties_national`, longitude = INTPTLONG)
`2015_Gaz_counties_national` <- `2015_Gaz_counties_national` %>% select(fips,latitude,longitude)
```

We then merged that dataset into our main dataframe, countyrates.

```{r}
countyrates <- left_join(countyrates,`2015_Gaz_counties_national`,by=c("fips"))
```

After combining the datasets, we noticed there were several observations with missing data for some variables. In particular, all of Alaska was missing most of the Deleetdk data and a few stray observations were missing the education data and a number of other variables. We removed those observations.

```{r}
countyratesnogaps <- countyrates %>% filter(fips>100) %>% filter(fips<70000) %>% filter(!is.na(rep16_frac)) %>% filter(!is.na(At.Least.High.School.Diploma))  %>% arrange(fips)
```

We also chose to remove some of the variables that had a lot of missing data by selecting the others.

```{r echo=FALSE}
countyratesnogaps <- countyratesnogaps %>% select(location,fips,HIV.HIV80,HIV.HIV85,HIV.HIV90,HIV.HIV95,HIV.HIV00,HIV.HIV05,HIV.HIV10,HIV.HIV14,INF.INF80,INF.INF85,INF.INF90,INF.INF95,INF.INF00,INF.INF05,INF.INF10,INF.INF14,TROP.TROP80,TROP.TROP85,TROP.TROP90,TROP.TROP95,TROP.TROP00,TROP.TROP05,TROP.TROP10,TROP.TROP14,MAT.MAT80,MAT.MAT85,MAT.MAT90,MAT.MAT95,MAT.MAT00,MAT.MAT05,MAT.MAT10,MAT.MAT14,NEON.NEON80,NEON.NEON85,NEON.NEON90,NEON.NEON95,NEON.NEON00,NEON.NEON05,NEON.NEON10,NEON.NEON14,NUT.NUT80,NUT.NUT85,NUT.NUT90,NUT.NUT95,NUT.NUT00,NUT.NUT05,NUT.NUT10,NUT.NUT14,OTHC.OTHC80,OTHC.OTHC85,OTHC.OTHC90,OTHC.OTHC95,OTHC.OTHC00,OTHC.OTHC05,OTHC.OTHC10,OTHC.OTHC14,NEOP.NEOP80,NEOP.NEOP85,NEOP.NEOP90,NEOP.NEOP95,NEOP.NEOP00,NEOP.NEOP05,NEOP.NEOP10,NEOP.NEOP14,CHRON.CHRON80,CHRON.CHRON85,CHRON.CHRON90,CHRON.CHRON95,CHRON.CHRON00,CHRON.CHRON05,CHRON.CHRON10,CHRON.CHRON14,CIRR.CIRR80,CIRR.CIRR85,CIRR.CIRR90,CIRR.CIRR95,CIRR.CIRR00,CIRR.CIRR05,CIRR.CIRR10,CIRR.CIRR14,DIG.DIG80,DIG.DIG85,DIG.DIG90,DIG.DIG95,DIG.DIG00,DIG.DIG05,DIG.DIG10,DIG.DIG14,NEUR.NEUR80,NEUR.NEUR85,NEUR.NEUR90,NEUR.NEUR95,NEUR.NEUR00,NEUR.NEUR05,NEUR.NEUR10,NEUR.NEUR14,MENSUB.MENSUB80,MENSUB.MENSUB85,MENSUB.MENSUB90,MENSUB.MENSUB95,MENSUB.MENSUB00,MENSUB.MENSUB05,MENSUB.MENSUB10,MENSUB.MENSUB14,DIAB.DIAB80,DIAB.DIAB85,DIAB.DIAB90,DIAB.DIAB95,DIAB.DIAB00,DIAB.DIAB05,DIAB.DIAB10,DIAB.DIAB14,MUSC.MUSC80,MUSC.MUSC85,MUSC.MUSC90,MUSC.MUSC95,MUSC.MUSC00,MUSC.MUSC05,MUSC.MUSC10,MUSC.MUSC14,OTHN.OTHN80,OTHN.OTHN85,OTHN.OTHN90,OTHN.OTHN95,OTHN.OTHN00,OTHN.OTHN05,OTHN.OTHN10,OTHN.OTHN14,TRAN.TRAN80,TRAN.TRAN85,TRAN.TRAN90,TRAN.TRAN95,TRAN.TRAN00,TRAN.TRAN05,TRAN.TRAN10,TRAN.TRAN14,UNIN.UNIN80,UNIN.UNIN85,UNIN.UNIN90,UNIN.UNIN95,UNIN.UNIN00,UNIN.UNIN05,UNIN.UNIN10,UNIN.UNIN14,SELF.SELF80,SELF.SELF85,SELF.SELF90,SELF.SELF95,SELF.SELF00,SELF.SELF05,SELF.SELF10,SELF.SELF14,WAR.WAR80,WAR.WAR85,WAR.WAR90,WAR.WAR95,WAR.WAR00,WAR.WAR05,WAR.WAR10,WAR.WAR14,rep16_frac,dem16_frac,Less.Than.High.School,At.Least.High.School.Diploma,At.Least.Bachelor.s.Degree,Graduate.Degree,School.Enrollment,Median.Earnings.2010.dollars,White.not.Latino.Population,African.American.Population,Native.American.Population,Asian.American.Population,Latino.Population,Adults.65.and.Older.Living.in.Poverty,Total.Population,Poverty.Rate.below.federal.poverty.threshold,Child.Poverty.living.in.families.below.the.poverty.line,Management.professional.and.related.occupations,Service.occupations,Sales.and.office.occupations,Farming.fishing.and.forestry.occupations,Construction.extraction.maintenance.and.repair.occupations,Production.transportation.and.material.moving.occupations,median_age,Children.in.single.parent.households,Adult.obesity,Diabetes,Uninsured,Unemployment,latitude,longitude)
```

Finally, we formally created the dataframe we would use of all the county observations with tidy data for all of our variables of interest.

```{r}
countyratesfullcases <- countyratesnogaps[complete.cases(countyratesnogaps),]
```


\newpage

**Model Creation/Statistical Computation**

We created two models for each of the twenty causes of death - a regression tree and a k nearest neighbors model.


```{r}
require(caret)
require(rpart)
require(rpart.plot)
library(tree)
```


To give a sense for our model-building process, we have included the R code for the model creation for the cause of death, HIV-AIDS and tuberculosis. All of the other models follow the same format, and only the regression trees and the mean square errors from the k-nearest neighbors model are shown for them.

*HIV: HIV-AIDS and Tuberculosis*

Regression tree and mean square error

```{r}
fips <- countyratesfullcases$fips
latitude <- countyratesfullcases$latitude
longitude <- countyratesfullcases$longitude
```

```{r}
require(dplyr)
set.seed(1)
fitControl <- trainControl(method="cv")
tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School 
+ At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree 
+ School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population 
+ African.American.Population + Native.American.Population + 
  Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty 
+ Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + 
Management.professional.and.related.occupations + Service.occupations + 
Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + 
Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + 
Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.HIV14$finalModel)

mortrate <- countyratesfullcases$HIV.HIV14
cause <- rep("HIV", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
HIVreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realHIV14 <- countyratesfullcases$HIV.HIV14
preds.tr.HIV14 <- predict(tr.HIV14$finalModel, newdata = countyratesfullcases)
HIV14 <- cbind.data.frame(preds.tr.HIV14,realHIV14)
HIV14 <- HIV14 %>% mutate(HIVtreediff = ((preds.tr.HIV14 - realHIV14)/realHIV14))
mortrate <- HIV14$HIVtreediff
cause <- rep("HIV", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
HIVtree <- cbind(fips,cause,model,mortrate) 

treeHIV14.mse <- mean((preds.tr.HIV14 - realHIV14)^2)
treeHIV14.mse
```


k-Nearest Neighbors mean square error


```{r}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.HIV14 <- train(HIV.HIV14 ~ rep16_frac + dem16_frac + Less.Than.High.School 
+ At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree 
+ School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population 
+ African.American.Population + Native.American.Population + 
  Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty 
+ Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + 
Management.professional.and.related.occupations + Service.occupations + 
Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + 
Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + 
Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))

preds.knn.tr.HIV14 <- predict(knn.tr.HIV14, newdata = countyratesfullcases)
realknnHIV14 <- countyratesfullcases$HIV.HIV14
knnHIV14 <- cbind.data.frame(preds.knn.tr.HIV14,realknnHIV14)
knnHIV14 <- knnHIV14 %>% mutate(HIVknndiff = ((preds.knn.tr.HIV14 - realknnHIV14)/realknnHIV14))
mortrate <- knnHIV14$HIVknndiff
cause <- rep("HIV", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
HIVknn <- cbind(fips,cause,model,mortrate)

knnHIV14.mse <- mean((preds.knn.tr.HIV14 - realknnHIV14)^2)
knnHIV14.mse
```


\newpage

*INF: Diarrhea, lower respiratory, and other common infectious diseases*

Regression tree and mean square error

```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.INF14 <- train(INF.INF14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.INF14$finalModel)
mortrate <- countyratesfullcases$INF.INF14
cause <- rep("INF", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
INFreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realINF14 <- countyratesfullcases$INF.INF14
preds.tr.INF14 <- predict(tr.INF14$finalModel, newdata = countyratesfullcases)
INF14 <- cbind.data.frame(preds.tr.INF14,realINF14)
INF14 <- INF14 %>% mutate(INFtreediff = ((preds.tr.INF14 - realINF14)/realINF14))
mortrate <- INF14$INFtreediff
cause <- rep("INF", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
INFtree <- cbind(fips,cause,model,mortrate) 

treeINF14.mse <- mean((preds.tr.INF14 - realINF14)^2)
treeINF14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.INF14 <- train(INF.INF14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.INF14 <- predict(knn.tr.INF14, newdata = countyratesfullcases)
realknnINF14 <- countyratesfullcases$INF.INF14
knnINF14 <- cbind.data.frame(preds.knn.tr.INF14,realknnINF14)
knnINF14 <- knnINF14 %>% mutate(INFknndiff = ((preds.knn.tr.INF14 - realknnINF14)/realknnINF14))
mortrate <- knnINF14$INFknndiff
cause <- rep("INF", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
INFknn <- cbind(fips,cause,model,mortrate)

knnINF14.mse <- mean((preds.knn.tr.INF14 - realknnINF14)^2)
knnINF14.mse
```


\newpage

*TROP: Neglected tropical diseases and malaria*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.TROP14 <- train(TROP.TROP14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
mortrate <- countyratesfullcases$TROP.TROP14
cause <- rep("TROP", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
TROPreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realTROP14 <- countyratesfullcases$TROP.TROP14
preds.tr.TROP14 <- predict(tr.TROP14$finalModel, newdata = countyratesfullcases)
TROP14 <- cbind.data.frame(preds.tr.TROP14,realTROP14)
TROP14 <- TROP14 %>% mutate(TROPtreediff = ((preds.tr.TROP14 - realTROP14)/realTROP14))
mortrate <- TROP14$TROPtreediff
cause <- rep("TROP", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
TROPtree <- cbind(fips,cause,model,mortrate) 

treeTROP14.mse <- mean((preds.tr.TROP14 - realTROP14)^2)
treeTROP14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.TROP14 <- train(TROP.TROP14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.TROP14 <- predict(knn.tr.TROP14, newdata = countyratesfullcases)
realknnTROP14 <- countyratesfullcases$TROP.TROP14
knnTROP14 <- cbind.data.frame(preds.knn.tr.TROP14,realknnTROP14)
knnTROP14 <- knnTROP14 %>% mutate(TROPknndiff = ((preds.knn.tr.TROP14 - realknnTROP14)/realknnTROP14))
mortrate <- knnTROP14$TROPknndiff
cause <- rep("TROP", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
TROPknn <- cbind(fips,cause,model,mortrate)

knnTROP14.mse <- mean((preds.knn.tr.TROP14 - realknnTROP14)^2)
knnTROP14.mse
```


\newpage

*MAT: Maternal disorders*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.MAT14 <- train(MAT.MAT14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.MAT14$finalModel)
mortrate <- countyratesfullcases$MAT.MAT14
cause <- rep("MAT", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
MATreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realMAT14 <- countyratesfullcases$MAT.MAT14
preds.tr.MAT14 <- predict(tr.MAT14$finalModel, newdata = countyratesfullcases)
MAT14 <- cbind.data.frame(preds.tr.MAT14,realMAT14)
MAT14 <- MAT14 %>% mutate(MATtreediff = ((preds.tr.MAT14 - realMAT14)/realMAT14))
mortrate <- MAT14$MATtreediff
cause <- rep("MAT", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
MATtree <- cbind(fips,cause,model,mortrate) 

treeMAT14.mse <- mean((preds.tr.MAT14 - realMAT14)^2)
treeMAT14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.MAT14 <- train(MAT.MAT14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.MAT14 <- predict(knn.tr.MAT14, newdata = countyratesfullcases)
realknnMAT14 <- countyratesfullcases$MAT.MAT14
knnMAT14 <- cbind.data.frame(preds.knn.tr.MAT14,realknnMAT14)
knnMAT14 <- knnMAT14 %>% mutate(MATknndiff = ((preds.knn.tr.MAT14 - realknnMAT14)/realknnMAT14))
mortrate <- knnMAT14$MATknndiff
cause <- rep("MAT", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
MATknn <- cbind(fips,cause,model,mortrate)

knnMAT14.mse <- mean((preds.knn.tr.MAT14 - realknnMAT14)^2)
knnMAT14.mse
```


\newpage

*NEON: Neonatal disorders*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.NEON14 <- train(NEON.NEON14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.NEON14$finalModel)
mortrate <- countyratesfullcases$NEON.NEON14
cause <- rep("NEON", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
NEONreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realNEON14 <- countyratesfullcases$NEON.NEON14
preds.tr.NEON14 <- predict(tr.NEON14$finalModel, newdata = countyratesfullcases)
NEON14 <- cbind.data.frame(preds.tr.NEON14,realNEON14)
NEON14 <- HIV14 %>% mutate(NEONtreediff = ((preds.tr.NEON14 - realNEON14)/realNEON14))
mortrate <- NEON14$NEONtreediff
cause <- rep("NEON", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
NEONtree <- cbind(fips,cause,model,mortrate) 

treeNEON14.mse <- mean((preds.tr.NEON14 - realNEON14)^2)
treeNEON14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.NEON14 <- train(NEON.NEON14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.NEON14 <- predict(knn.tr.NEON14, newdata = countyratesfullcases)
realknnNEON14 <- countyratesfullcases$NEON.NEON14
knnNEON14 <- cbind.data.frame(preds.knn.tr.NEON14,realknnNEON14)
knnNEON14 <- knnNEON14 %>% mutate(NEONknndiff = ((preds.knn.tr.NEON14 - realknnNEON14)/realknnNEON14))
mortrate <- knnNEON14$NEONknndiff
cause <- rep("NEON", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
NEONknn <- cbind(fips,cause,model,mortrate)

knnNEON14.mse <- mean((preds.knn.tr.NEON14 - realknnNEON14)^2)
knnNEON14.mse
```


\newpage

*NUT: Nutritional deficiencies*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.NUT14 <- train(NUT.NUT14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.NUT14$finalModel)
mortrate <- countyratesfullcases$NUT.NUT14
cause <- rep("NUT", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
NUTreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realNUT14 <- countyratesfullcases$NUT.NUT14
preds.tr.NUT14 <- predict(tr.NUT14$finalModel, newdata = countyratesfullcases)
NUT14 <- cbind.data.frame(preds.tr.NUT14,realNUT14)
NUT14 <- NUT14 %>% mutate(NUTtreediff = ((preds.tr.NUT14 - realNUT14)/realNUT14))
mortrate <- NUT14$NUTtreediff
cause <- rep("NUT", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
NUTtree <- cbind(fips,cause,model,mortrate) 

treeNUT14.mse <- mean((preds.tr.NUT14 - realNUT14)^2)
treeNUT14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.NUT14 <- train(NUT.NUT14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.NUT14 <- predict(knn.tr.NUT14, newdata = countyratesfullcases)
realknnNUT14 <- countyratesfullcases$NUT.NUT14
knnNUT14 <- cbind.data.frame(preds.knn.tr.NUT14,realknnNUT14)
knnNUT14 <- knnNUT14 %>% mutate(NUTknndiff = ((preds.knn.tr.NUT14 - realknnNUT14)/realknnNUT14))
mortrate <- knnNUT14$NUTknndiff
cause <- rep("NUT", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
NUTknn <- cbind(fips,cause,model,mortrate)

knnNUT14.mse <- mean((preds.knn.tr.NUT14 - realknnNUT14)^2)
knnNUT14.mse
```


\newpage

*OTHC: Other communicable, maternal, neonatal, and nutritional diseases*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.OTHC14 <- train(OTHC.OTHC14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
mortrate <- countyratesfullcases$OTHC.OTHC14
cause <- rep("OTHC", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
OTHCreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realOTHC14 <- countyratesfullcases$OTHC.OTHC14
preds.tr.OTHC14 <- predict(tr.OTHC14$finalModel, newdata = countyratesfullcases)
OTHC14 <- cbind.data.frame(preds.tr.OTHC14,realOTHC14)
OTHC14 <- HIV14 %>% mutate(OTHCtreediff = ((preds.tr.OTHC14 - realOTHC14)/realOTHC14))
mortrate <- OTHC14$OTHCtreediff
cause <- rep("OTHC", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
OTHCtree <- cbind(fips,cause,model,mortrate) 

treeOTHC14.mse <- mean((preds.tr.OTHC14 - realOTHC14)^2)
treeOTHC14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.OTHC14 <- train(OTHC.OTHC14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.OTHC14 <- predict(knn.tr.OTHC14, newdata = countyratesfullcases)
realknnOTHC14 <- countyratesfullcases$OTHC.OTHC14
knnOTHC14 <- cbind.data.frame(preds.knn.tr.OTHC14,realknnOTHC14)
knnOTHC14 <- knnOTHC14 %>% mutate(OTHCknndiff = ((preds.knn.tr.OTHC14 - realknnOTHC14)/realknnOTHC14))
mortrate <- knnOTHC14$OTHCknndiff
cause <- rep("OTHC", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
OTHCknn <- cbind(fips,cause,model,mortrate)

knnOTHC14.mse <- mean((preds.knn.tr.OTHC14 - realknnOTHC14)^2)
knnOTHC14.mse
```


\newpage

*NEOP: Neoplasms*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.NEOP14 <- train(NEOP.NEOP14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.NEOP14$finalModel)
mortrate <- countyratesfullcases$NEOP.NEOP14
cause <- rep("NEOP", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
NEOPreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realNEOP14 <- countyratesfullcases$NEOP.NEOP14
preds.tr.NEOP14 <- predict(tr.NEOP14$finalModel, newdata = countyratesfullcases)
NEOP14 <- cbind.data.frame(preds.tr.NEOP14,realNEOP14)
NEOP14 <- NEOP14 %>% mutate(NEOPtreediff = ((preds.tr.NEOP14 - realNEOP14)/realNEOP14))
mortrate <- NEOP14$NEOPtreediff
cause <- rep("NEOP", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
NEOPtree <- cbind(fips,cause,model,mortrate) 

treeNEOP14.mse <- mean((preds.tr.NEOP14 - realNEOP14)^2)
treeNEOP14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.NEOP14 <- train(NEOP.NEOP14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.NEOP14 <- predict(knn.tr.NEOP14, newdata = countyratesfullcases)
realknnNEOP14 <- countyratesfullcases$NEOP.NEOP14
knnNEOP14 <- cbind.data.frame(preds.knn.tr.NEOP14,realknnNEOP14)
knnNEOP14 <- knnHIV14 %>% mutate(NEOPknndiff = ((preds.knn.tr.NEOP14 - realknnNEOP14)/realknnNEOP14))
mortrate <- knnNEOP14$NEOPknndiff
cause <- rep("NEOP", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
NEOPknn <- cbind(fips,cause,model,mortrate)

knnNEOP14.mse <- mean((preds.knn.tr.NEOP14 - realknnNEOP14)^2)
knnNEOP14.mse
```


\newpage

*CHRON: Chronic respiratory diseases*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.CHRON14 <- train(CHRON.CHRON14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.CHRON14$finalModel)
mortrate <- countyratesfullcases$CHRON.CHRON14
cause <- rep("CHRON", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
CHRONreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realCHRON14 <- countyratesfullcases$CHRON.CHRON14
preds.tr.CHRON14 <- predict(tr.CHRON14$finalModel, newdata = countyratesfullcases)
CHRON14 <- cbind.data.frame(preds.tr.CHRON14,realCHRON14)
CHRON14 <- CHRON14 %>% mutate(CHRONtreediff = ((preds.tr.CHRON14 - realCHRON14)/realCHRON14))
mortrate <- CHRON14$CHRONtreediff
cause <- rep("CHRON", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
CHRONtree <- cbind(fips,cause,model,mortrate) 

treeCHRON14.mse <- mean((preds.tr.CHRON14 - realCHRON14)^2)
treeCHRON14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.CHRON14 <- train(CHRON.CHRON14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.CHRON14 <- predict(knn.tr.CHRON14, newdata = countyratesfullcases)
realknnCHRON14 <- countyratesfullcases$CHRON.CHRON14
knnCHRON14 <- cbind.data.frame(preds.knn.tr.CHRON14,realknnCHRON14)
knnCHRON14 <- knnCHRON14 %>% mutate(CHRONknndiff = ((preds.knn.tr.CHRON14 - realknnCHRON14)/realknnCHRON14))
mortrate <- knnCHRON14$CHRONknndiff
cause <- rep("CHRON", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
CHRONknn <- cbind(fips,cause,model,mortrate)

knnCHRON14.mse <- mean((preds.knn.tr.CHRON14 - realknnCHRON14)^2)
knnCHRON14.mse
```


\newpage

*CIRR: Cirrhosis and other chronic liver diseases*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.CIRR14 <- train(CIRR.CIRR14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.CIRR14$finalModel)
mortrate <- countyratesfullcases$CIRR.CIRR14
cause <- rep("CIRR", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
CIRRreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realCIRR14 <- countyratesfullcases$CIRR.CIRR14
preds.tr.CIRR14 <- predict(tr.CIRR14$finalModel, newdata = countyratesfullcases)
CIRR14 <- cbind.data.frame(preds.tr.CIRR14,realCIRR14)
CIRR14 <- CIRR14 %>% mutate(CIRRtreediff = ((preds.tr.CIRR14 - realCIRR14)/realCIRR14))
mortrate <- CIRR14$CIRRtreediff
cause <- rep("CIRR", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
CIRRtree <- cbind(fips,cause,model,mortrate) 

treeCIRR14.mse <- mean((preds.tr.CIRR14 - realCIRR14)^2)
treeCIRR14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.CIRR14 <- train(CIRR.CIRR14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.CIRR14 <- predict(knn.tr.CIRR14, newdata = countyratesfullcases)
realknnCIRR14 <- countyratesfullcases$CIRR.CIRR14
knnCIRR14 <- cbind.data.frame(preds.knn.tr.CIRR14,realknnCIRR14)
knnCIRR14 <- knnCIRR14 %>% mutate(CIRRknndiff = ((preds.knn.tr.CIRR14 - realknnCIRR14)/realknnCIRR14))
mortrate <- knnCIRR14$CIRRknndiff
cause <- rep("CIRR", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
CIRRknn <- cbind(fips,cause,model,mortrate)

knnCIRR14.mse <- mean((preds.knn.tr.CIRR14 - realknnCIRR14)^2)
knnCIRR14.mse
```


\newpage

*DIG: Digestive diseases*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.DIG14 <- train(DIG.DIG14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.DIG14$finalModel)
mortrate <- countyratesfullcases$DIG.DIG14
cause <- rep("DIG", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
DIGreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realDIG14 <- countyratesfullcases$DIG.DIG14
preds.tr.DIG14 <- predict(tr.DIG14$finalModel, newdata = countyratesfullcases)
DIG14 <- cbind.data.frame(preds.tr.DIG14,realDIG14)
DIG14 <- DIG14 %>% mutate(DIGtreediff = ((preds.tr.DIG14 - realDIG14)/realDIG14))
mortrate <- DIG14$DIGtreediff
cause <- rep("DIG", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
DIGtree <- cbind(fips,cause,model,mortrate) 

treeDIG14.mse <- mean((preds.tr.DIG14 - realDIG14)^2)
treeDIG14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.DIG14 <- train(DIG.DIG14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.DIG14 <- predict(knn.tr.DIG14, newdata = countyratesfullcases)
realknnDIG14 <- countyratesfullcases$DIG.DIG14
knnDIG14 <- cbind.data.frame(preds.knn.tr.DIG14,realknnDIG14)
knnDIG14 <- knnDIG14 %>% mutate(DIGknndiff = ((preds.knn.tr.DIG14 - realknnDIG14)/realknnDIG14))
mortrate <- knnDIG14$DIGknndiff
cause <- rep("DIG", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
DIGknn <- cbind(fips,cause,model,mortrate)

knnDIG14.mse <- mean((preds.knn.tr.DIG14 - realknnDIG14)^2)
knnDIG14.mse
```


\newpage

*NEUR: Neurological disorders*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.NEUR14 <- train(NEUR.NEUR14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.NEUR14$finalModel)
mortrate <- countyratesfullcases$NEUR.NEUR14
cause <- rep("NEUR", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
NEURreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realNEUR14 <- countyratesfullcases$NEUR.NEUR14
preds.tr.NEUR14 <- predict(tr.NEUR14$finalModel, newdata = countyratesfullcases)
NEUR14 <- cbind.data.frame(preds.tr.NEUR14,realNEUR14)
NEUR14 <- NEUR14 %>% mutate(NEURtreediff = ((preds.tr.NEUR14 - realNEUR14)/realNEUR14))
mortrate <- NEUR14$NEURtreediff
cause <- rep("NEUR", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
NEURtree <- cbind(fips,cause,model,mortrate) 

treeNEUR14.mse <- mean((preds.tr.NEUR14 - realNEUR14)^2)
treeNEUR14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.NEUR14 <- train(NEUR.NEUR14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.NEUR14 <- predict(knn.tr.NEUR14, newdata = countyratesfullcases)
realknnNEUR14 <- countyratesfullcases$NEUR.NEUR14
knnNEUR14 <- cbind.data.frame(preds.knn.tr.NEUR14,realknnNEUR14)
knnNEUR14 <- knnNEUR14 %>% mutate(NEURknndiff = ((preds.knn.tr.NEUR14 - realknnNEUR14)/realknnNEUR14))
mortrate <- knnNEUR14$NEURknndiff
cause <- rep("NEUR", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
NEURknn <- cbind(fips,cause,model,mortrate)

knnNEUR14.mse <- mean((preds.knn.tr.NEUR14 - realknnNEUR14)^2)
knnNEUR14.mse
```


\newpage

*MENSUB: Mental and substance use disorders*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.MENSUB14 <- train(MENSUB.MENSUB14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.MENSUB14$finalModel)
mortrate <- countyratesfullcases$MENSUB.MENSUB14
cause <- rep("MENSUB", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
MENSUBreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realMENSUB14 <- countyratesfullcases$MENSUB.MENSUB14
preds.tr.MENSUB14 <- predict(tr.MENSUB14$finalModel, newdata = countyratesfullcases)
MENSUB14 <- cbind.data.frame(preds.tr.MENSUB14,realMENSUB14)
MENSUB14 <- MENSUB14 %>% mutate(MENSUBtreediff = ((preds.tr.MENSUB14 - realMENSUB14)/realMENSUB14))
mortrate <- MENSUB14$MENSUBtreediff
cause <- rep("MENSUB", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
MENSUBtree <- cbind(fips,cause,model,mortrate) 

treeMENSUB14.mse <- mean((preds.tr.MENSUB14 - realMENSUB14)^2)
treeMENSUB14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.MENSUB14 <- train(MENSUB.MENSUB14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.MENSUB14 <- predict(knn.tr.MENSUB14, newdata = countyratesfullcases)
realknnMENSUB14 <- countyratesfullcases$MENSUB.MENSUB14
knnMENSUB14 <- cbind.data.frame(preds.knn.tr.MENSUB14,realknnMENSUB14)
knnMENSUB14 <- knnHIV14 %>% mutate(MENSUBknndiff = ((preds.knn.tr.MENSUB14 - realknnMENSUB14)/realknnMENSUB14))
mortrate <- knnMENSUB14$MENSUBknndiff
cause <- rep("MENSUB", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
MENSUBknn <- cbind(fips,cause,model,mortrate)

knnMENSUB14.mse <- mean((preds.knn.tr.MENSUB14 - realknnMENSUB14)^2)
knnMENSUB14.mse
```


\newpage

*DIAB: Diabetes, urogenital, blood, and endocrine diseases*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.DIAB14 <- train(DIAB.DIAB14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.DIAB14$finalModel)
mortrate <- countyratesfullcases$DIAB.DIAB14
cause <- rep("DIAB", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
DIABreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realDIAB14 <- countyratesfullcases$DIAB.DIAB14
preds.tr.DIAB14 <- predict(tr.DIAB14$finalModel, newdata = countyratesfullcases)
DIAB14 <- cbind.data.frame(preds.tr.DIAB14,realDIAB14)
DIAB14 <- HIV14 %>% mutate(DIABtreediff = ((preds.tr.DIAB14 - realDIAB14)/realDIAB14))
mortrate <- DIAB14$HIVtreediff
cause <- rep("DIAB", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
DIABtree <- cbind(fips,cause,model,mortrate) 

treeDIAB14.mse <- mean((preds.tr.DIAB14 - realDIAB14)^2)
treeDIAB14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.DIAB14 <- train(DIAB.DIAB14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.DIAB14 <- predict(knn.tr.DIAB14, newdata = countyratesfullcases)
realknnDIAB14 <- countyratesfullcases$DIAB.DIAB14
knnDIAB14 <- cbind.data.frame(preds.knn.tr.DIAB14,realknnDIAB14)
knnDIAB14 <- knnDIAB14 %>% mutate(DIABknndiff = ((preds.knn.tr.DIAB14 - realknnDIAB14)/realknnDIAB14))
mortrate <- knnDIAB14$DIABknndiff
cause <- rep("DIAB", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
DIABknn <- cbind(fips,cause,model,mortrate)

knnDIAB14.mse <- mean((preds.knn.tr.DIAB14 - realknnDIAB14)^2)
knnDIAB14.mse
```


\newpage

*MUSC: Musculoskeletal disorders*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.MUSC14 <- train(MUSC.MUSC14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.MUSC14$finalModel)
mortrate <- countyratesfullcases$MUSC.MUSC14
cause <- rep("MUSC", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
MUSCreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realMUSC14 <- countyratesfullcases$MUSC.MUSC14
preds.tr.MUSC14 <- predict(tr.MUSC14$finalModel, newdata = countyratesfullcases)
MUSC14 <- cbind.data.frame(preds.tr.MUSC14,realMUSC14)
MUSC14 <- HIV14 %>% mutate(MUSCtreediff = ((preds.tr.MUSC14 - realMUSC14)/realMUSC14))
mortrate <- MUSC14$MUSCtreediff
cause <- rep("MUSC", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
MUSCtree <- cbind(fips,cause,model,mortrate) 

treeMUSC14.mse <- mean((preds.tr.MUSC14 - realMUSC14)^2)
treeMUSC14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.MUSC14 <- train(MUSC.MUSC14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.MUSC14 <- predict(knn.tr.MUSC14, newdata = countyratesfullcases)
realknnMUSC14 <- countyratesfullcases$MUSC.MUSC14
knnMUSC14 <- cbind.data.frame(preds.knn.tr.MUSC14,realknnMUSC14)
knnMUSC14 <- knnMUSC14 %>% mutate(MUSCknndiff = ((preds.knn.tr.MUSC14 - realknnMUSC14)/realknnMUSC14))
mortrate <- knnMUSC14$MUSCknndiff
cause <- rep("MUSC", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
MUSCknn <- cbind(fips,cause,model,mortrate)

knnMUSC14.mse <- mean((preds.knn.tr.MUSC14 - realknnMUSC14)^2)
knnMUSC14.mse
```


\newpage

*OTHN: Other non-communicable diseases*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.OTHN14 <- train(OTHN.OTHN14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.OTHN14$finalModel)
mortrate <- countyratesfullcases$OTHN.OTHN14
cause <- rep("OTHN", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
OTHNreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realOTHN14 <- countyratesfullcases$OTHN.OTHN14
preds.tr.OTHN14 <- predict(tr.OTHN14$finalModel, newdata = countyratesfullcases)
OTHN14 <- cbind.data.frame(preds.tr.OTHN14,realOTHN14)
OTHN14 <- OTHN14 %>% mutate(OTHNtreediff = ((preds.tr.OTHN14 - realOTHN14)/realOTHN14))
mortrate <- OTHN14$OTHNtreediff
cause <- rep("OTHN", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
OTHNtree <- cbind(fips,cause,model,mortrate) 

treeOTHN14.mse <- mean((preds.tr.OTHN14 - realOTHN14)^2)
treeOTHN14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.OTHN14 <- train(OTHN.OTHN14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.OTHN14 <- predict(knn.tr.OTHN14, newdata = countyratesfullcases)
realknnOTHN14 <- countyratesfullcases$OTHN.OTHN14
knnOTHN14 <- cbind.data.frame(preds.knn.tr.OTHN14,realknnOTHN14)
knnOTHN14 <- knnOTHN14 %>% mutate(OTHNknndiff = ((preds.knn.tr.OTHN14 - realknnOTHN14)/realknnOTHN14))
mortrate <- knnOTHN14$OTHNknndiff
cause <- rep("OTHN", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
OTHNknn <- cbind(fips,cause,model,mortrate)

knnOTHN14.mse <- mean((preds.knn.tr.OTHN14 - realknnOTHN14)^2)
knnOTHN14.mse
```


\newpage

*TRAN: Transport injuries*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.TRAN14 <- train(TRAN.TRAN14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.TRAN14$finalModel)
mortrate <- countyratesfullcases$TRAN.TRAN14
cause <- rep("TRAN", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
TRANreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realTRAN14 <- countyratesfullcases$TRAN.TRAN14
preds.tr.TRAN14 <- predict(tr.TRAN14$finalModel, newdata = countyratesfullcases)
TRAN14 <- cbind.data.frame(preds.tr.TRAN14,realTRAN14)
TRAN14 <- TRAN14 %>% mutate(TRANtreediff = ((preds.tr.TRAN14 - realTRAN14)/realTRAN14))
mortrate <- TRAN14$TRANtreediff
cause <- rep("TRAN", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
TRANtree <- cbind(fips,cause,model,mortrate) 

treeTRAN14.mse <- mean((preds.tr.TRAN14 - realTRAN14)^2)
treeTRAN14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.TRAN14 <- train(TRAN.TRAN14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.TRAN14 <- predict(knn.tr.TRAN14, newdata = countyratesfullcases)
realknnTRAN14 <- countyratesfullcases$TRAN.TRAN14
knnTRAN14 <- cbind.data.frame(preds.knn.tr.TRAN14,realknnTRAN14)
knnTRAN14 <- knnTRAN14 %>% mutate(TRANknndiff = ((preds.knn.tr.TRAN14 - realknnTRAN14)/realknnTRAN14))
mortrate <- knnTRAN14$TRANknndiff
cause <- rep("TRAN", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
TRANknn <- cbind(fips,cause,model,mortrate)

knnTRAN14.mse <- mean((preds.knn.tr.TRAN14 - realknnTRAN14)^2)
knnTRAN14.mse
```


\newpage

*UNIN: Unintentional injuries*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.UNIN14 <- train(UNIN.UNIN14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.UNIN14$finalModel)
mortrate <- countyratesfullcases$UNIN.UNIN14
cause <- rep("UNIN", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
UNINreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realUNIN14 <- countyratesfullcases$UNIN.UNIN14
preds.tr.UNIN14 <- predict(tr.UNIN14$finalModel, newdata = countyratesfullcases)
UNIN14 <- cbind.data.frame(preds.tr.UNIN14,realUNIN14)
UNIN14 <- UNIN14 %>% mutate(UNINtreediff = ((preds.tr.UNIN14 - realUNIN14)/realUNIN14))
mortrate <- UNIN14$UNINtreediff
cause <- rep("UNIN", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
UNINtree <- cbind(fips,cause,model,mortrate) 

treeUNIN14.mse <- mean((preds.tr.UNIN14 - realUNIN14)^2)
treeUNIN14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.UNIN14 <- train(UNIN.UNIN14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.UNIN14 <- predict(knn.tr.UNIN14, newdata = countyratesfullcases)
realknnUNIN14 <- countyratesfullcases$UNIN.UNIN14
knnUNIN14 <- cbind.data.frame(preds.knn.tr.UNIN14,realknnUNIN14)
knnUNIN14 <- knnUNIN14 %>% mutate(UNINknndiff = ((preds.knn.tr.UNIN14 - realknnUNIN14)/realknnUNIN14))
mortrate <- knnUNIN14$UNINknndiff
cause <- rep("UNIN", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
UNINknn <- cbind(fips,cause,model,mortrate)

knnUNIN14.mse <- mean((preds.knn.tr.UNIN14 - realknnUNIN14)^2)
knnUNIN14.mse
```


\newpage

*SELF: Self-harm and interpersonal violence*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.SELF14 <- train(SELF.SELF14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.SELF14$finalModel)
mortrate <- countyratesfullcases$SELF.SELF14
cause <- rep("SELF", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
SELFreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realSELF14 <- countyratesfullcases$SELF.SELF14
preds.tr.SELF14 <- predict(tr.SELF14$finalModel, newdata = countyratesfullcases)
SELF14 <- cbind.data.frame(preds.tr.SELF14,realSELF14)
SELF14 <- SELF14 %>% mutate(SELFtreediff = ((preds.tr.SELF14 - realSELF14)/realSELF14))
mortrate <- SELF14$SELFtreediff
cause <- rep("SELF", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
SELFtree <- cbind(fips,cause,model,mortrate) 

treeSELF14.mse <- mean((preds.tr.SELF14 - realSELF14)^2)
treeSELF14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.SELF14 <- train(SELF.SELF14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.SELF14 <- predict(knn.tr.SELF14, newdata = countyratesfullcases)
realknnSELF14 <- countyratesfullcases$SELF.SELF14
knnSELF14 <- cbind.data.frame(preds.knn.tr.SELF14,realknnSELF14)
knnSELF14 <- knnSELF14 %>% mutate(SELFknndiff = ((preds.knn.tr.SELF14 - realknnSELF14)/realknnSELF14))
mortrate <- knnSELF14$SELFknndiff
cause <- rep("SELF", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
SELFknn <- cbind(fips,cause,model,mortrate)

knnSELF14.mse <- mean((preds.knn.tr.SELF14 - realknnSELF14)^2)
knnSELF14.mse
```


\newpage

*WAR: Forces of nature, war, and legal intervention*

Regression tree and mean square error


```{r echo=FALSE}
set.seed(1)
fitControl <- trainControl(method="cv")
tr.WAR14 <- train(WAR.WAR14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, data=countyratesfullcases, method="rpart2", 
                  trControl = fitControl, tuneGrid=data.frame(maxdepth=1:20))
rpart.plot(tr.WAR14$finalModel)
mortrate <- countyratesfullcases$WAR.WAR14
cause <- rep("WAR", 3110)
cause <- as.data.frame(cause)
model <- rep("real", 3110)
model <- as.data.frame(model)
WARreal <- cbind(fips,cause,model,mortrate)
```

```{r}
realWAR14 <- countyratesfullcases$WAR.WAR14
preds.tr.WAR14 <- predict(tr.WAR14$finalModel, newdata = countyratesfullcases)
WAR14 <- cbind.data.frame(preds.tr.WAR14,realWAR14)
WAR14 <- WAR14 %>% mutate(WARtreediff = ((preds.tr.WAR14 - realWAR14)/realWAR14))
mortrate <- WAR14$WARtreediff
cause <- rep("WAR", 3110)
cause <- as.data.frame(cause)
model <- rep("tree", 3110)
model <- as.data.frame(model)
WARtree <- cbind(fips,cause,model,mortrate) 

treeWAR14.mse <- mean((preds.tr.WAR14 - realWAR14)^2)
treeWAR14.mse
```


k-Nearest Neighbors mean square error


```{r echo=FALSE}
set.seed(4747)
fitControl <- trainControl(method="none")
knn.tr.WAR14 <- train(WAR.WAR14 ~ rep16_frac + dem16_frac + Less.Than.High.School + At.Least.High.School.Diploma + At.Least.Bachelor.s.Degree + Graduate.Degree + School.Enrollment + Median.Earnings.2010.dollars + White.not.Latino.Population + African.American.Population + Native.American.Population + Asian.American.Population + Latino.Population + Adults.65.and.Older.Living.in.Poverty + Total.Population + Poverty.Rate.below.federal.poverty.threshold + Child.Poverty.living.in.families.below.the.poverty.line + Management.professional.and.related.occupations +      Service.occupations + Sales.and.office.occupations + Farming.fishing.and.forestry.occupations + Construction.extraction.maintenance.and.repair.occupations + Production.transportation.and.material.moving.occupations + median_age + Children.in.single.parent.households + Adult.obesity + Diabetes + Uninsured + Unemployment, 
                data=countyratesfullcases, method="knn", trControl = fitControl, 
                 tuneGrid= data.frame(k=5))
preds.knn.tr.WAR14 <- predict(knn.tr.WAR14, newdata = countyratesfullcases)
realknnWAR14 <- countyratesfullcases$WAR.WAR14
knnWAR14 <- cbind.data.frame(preds.knn.tr.WAR14,realknnWAR14)
knnWAR14 <- knnWAR14 %>% mutate(WARknndiff = ((preds.knn.tr.WAR14 - realknnWAR14)/realknnWAR14))
mortrate <- knnWAR14$WARknndiff
cause <- rep("WAR", 3110)
cause <- as.data.frame(cause)
model <- rep("knn", 3110)
model <- as.data.frame(model)
WARknn <- cbind(fips,cause,model,mortrate)

knnWAR14.mse <- mean((preds.knn.tr.WAR14 - realknnWAR14)^2)
knnWAR14.mse
```


**Shiny Visualization**

The first important step to building our shiny application was to clean the data in a way that would be "function friendly." While researching Shiny and finding ways to build interactive maps of US county data, we discovered a package called choroplethr. From this package we utilize a function called county_choropleth() which builds a choropleth map of the counties in the United States. However, count-choropleth() takes a datafram with specific column names: region and value. Region must be assigned to the standard fips codes for the United States counties without any leading zeros. Value is then the information that will be projected onto the map. By setting number of colors to 1 in this instance we are able to create a continuous scale for our variable of interest - mortality rates. However, this particular function that county-choropleth() takes means that we need to manipulate our data set in several ways. This required us to rename our fips column to region and our mortality colu...

```{r echo =FALSE}
fips <- countyratesfullcases$fips
results  <- rbind.data.frame(
      HIVreal,HIVtree,HIVknn,
      INFreal,INFtree,INFknn,
      TROPreal,TROPtree,TROPknn,
      MATreal,MATtree,MATknn,
      NEONreal,NEONtree,NEONknn,
      NUTreal,NUTtree,NUTknn,
      OTHCreal,OTHCtree,OTHCknn,
      NEOPreal,NEOPtree,NEOPknn,
      CHRONreal,CHRONtree,CHRONknn,
      CIRRreal,CIRRtree,CIRRknn,
      DIGreal,DIGtree,DIGknn,
      NEURreal,NEURtree,NEURknn,
      MENSUBreal,MENSUBtree,MENSUBknn,
      DIABreal,DIABtree,DIABknn,
      MUSCreal,MUSCtree,MUSCknn,
      OTHNreal,OTHNtree,OTHNknn,
      TRANreal,TRANtree,TRANknn,
      UNINreal,UNINtree,UNINknn,
      SELFreal,SELFtree,SELFknn,
      WARreal,WARtree,WARknn)
View(results)

write.csv(results,"results.csv")
```

**Limitations**

